---
title: "Assignment B-1: Making a function"
output: github_document
---

This assignment covers making a function in R, documenting it, and testing it.

## Getting Started

Load the requirements for this assignment:

```{r}
library(palmerpenguins)
library(gapminder)
library(tidyverse)
library(testthat)
```

## Exercise 1 & 2: Make a Function &  Document your Function

In this exercise, I’ll make a function, fortify it, and document the function using roxygen2 tags.

Notice that in the return statement, we used `as_tibble` to ensure the output type is consistent despite the type of input.

```{r}
#' Count of missing values
#'
#' @description `count_na` returns the count of missing values(NA) in the specified column `col_name` of the given data set `data`
#'
#' @param data data set to get the specific column from
#' @param col_name name of the column (string input) to count missing values.
#'
#' @return A 1 x 1 tibble with variable "col_name" and its value "integer count of missing values"
count_na <- function(data, col_name) {
  if (!is.character(col_name)) {
    stop("col_name should be a string.")
  } else if (!(col_name %in% colnames(data))) {
    stop("'", col_name, "' does not exist in the data.")
  } else {
    return(as_tibble(data %>% summarise(across(col_name, ~ sum(is.na(.x))))))
  }
}
```
## Exercise 3: Include examples

First, we need to check if we have the data set we want to test with. We're gonna use `penguins` data set from `palmerpenguins` package.

```{r}
head(penguins)
```

For example, I would like to see how many missing values are there in `sex`, `bill_length_mm`, and `year` variables in data set `penguins`, respectively.

```{r}
count_na(penguins, "sex")
count_na(penguins, "bill_length_mm")
count_na(penguins, "year")
```

To show the error messages this function deliberately throws, we can look at the following two examples:

1. We pass in a column name `col_name` that is not a string. It will throw error with message `col_name should be a string.`

2. We pass in a column name `col_name` that does not exist in the data. It will give error with message `'{col_name}' does not exist in the data.`

```{r, error=TRUE}
count_na(penguins, 22)
count_na(penguins, "month")
```

## Exercise 4: Test the Function

The first three tests are testing if the function gives the expected output. The last three tests are testing whether the function throws the expected errors. The output and errors are similar to the examples shown above.

```{r}
test_that("Test the function 'count_na' with data set 'penguins'", {
  expect_equal(count_na(penguins, "sex"), as_tibble(summarise(penguins, across(sex, ~ sum(is.na(.x))))))
  expect_equal(count_na(penguins, "bill_length_mm"), as_tibble(summarise(penguins, across(bill_length_mm, ~ sum(is.na(.x))))))
  expect_equal(count_na(penguins, "year"), as_tibble(summarise(penguins, across(year, ~ sum(is.na(.x))))))
  expect_error(count_na(penguins, 22), "col_name should be a string.")
  expect_error(count_na(penguins, ""), "'' does not exist in the data.")
  expect_error(count_na(penguins, "month"), "'month' does not exist in the data.")
})
```

The following test cases test the function with the data set `mtcars`, which is a data frame. We can then use these test cases to see if the output is consistent despite the input is in different types.

```{r}
head(mtcars)
```

```{r}
test_that("Test the function 'count_na' with data set 'mtcars'", {
  expect_equal(count_na(mtcars, "cyl"), as_tibble(summarise(mtcars, across(cyl, ~ sum(is.na(.x))))))
  expect_equal(count_na(mtcars, "disp"), as_tibble(summarise(mtcars, across(disp, ~ sum(is.na(.x))))))
  expect_equal(count_na(mtcars, "vs"), as_tibble(summarise(mtcars, across(vs, ~ sum(is.na(.x))))))
  expect_error(count_na(mtcars, 22), "col_name should be a string.")
  expect_error(count_na(mtcars, ""), "'' does not exist in the data.")
  expect_error(count_na(mtcars, "month"), "'month' does not exist in the data.")
})
```
